# Intro
This code was originally worked on by Taz Saad, and then cleaned, compiled, and edited by Aziz Zafar. Below are some functions that perform different statistical tests. Out of these, chunk 11 runs most of these tests using a nested for loop. To select the tests to be used, or the outcome variables to be used, you can modify only chunk 11 accordingly. Majority of the libraries to be used are loaded in chunk 1. Contact @azafar@colgate.edu for any queries or follow-up questions.



# Univariate Logistic Regression

```{r, results='hide', include = FALSE}
library(tidyverse)
library(bestglm)
library(foreign)
library(descr)
library(dplyr)
univariate_lr <- function(gender, disease) {
  options(scipen=999)
  print(disease)
  df <- data.frame(Variable_Name=character(), Odds_Ratio=numeric(), Conf_25=numeric(), Conf_75=numeric(), P_Value=numeric())
  f_path = paste(disease,'/split/',gender,sep='')
  files <- list.files(path=f_path, pattern="*.sav", full.names=TRUE, recursive=FALSE)

  for (f in files) {
    data <- read.spss(file =f, to.data.frame = T)
    names(data)[ncol(data)] <- 'y'
    res = glm(y ~ .,data=data,family=binomial())
    trm <- attr(res$terms, "term.labels")
    for (i in (seq(2,dim(data)[2],by=1))) {
      res_vec <- tryCatch(c(trm[i-1], exp(summary(res)$coef[i,][1]), exp(confint(res)[i,]), summary(res)$coef[i,][4]), error=function(e) c(NA,NA,NA,NA,NA))
      df[nrow(df) + 1,] = res_vec
    }
    
  }
  adjusted <- p.adjust(df[, c('P_Value')], method="BH")
  df$Adjusted_P = adjusted
  
  df <- df[order(df$P_Value),]
  df
}
```

```{r, results='hide', include = FALSE}
univariate_lr_top <- function(disease) {
  options(scipen=999)
  f_path = paste(disease, ".csv", sep ="")
  data <- data.frame(read.csv(f_path))
  data = data[,!names(data) %in% c("X")]
  
  df <- data.frame(Variable_Name=character(), Odds_Ratio=numeric(), Conf_25=numeric(), Conf_75=numeric(), P_Value=numeric())
  
  if (grepl('BDI',disease,fixed=TRUE)) {
    data = data %>% dplyr::select(-BDI,BDI)
  } else if (grepl('STAI',disease,fixed=TRUE)) {
    data = data %>% dplyr::select(-STAI,STAI)
  } else if (grepl('SPAQ',disease,fixed=TRUE)){
    data = data %>% dplyr::select(-SPAQ,SPAQ)
  } else if (grepl('SLEEP',disease,fixed=TRUE)){
    data = data %>% dplyr::select(-SLEEP,SLEEP)
  } else {
    data = data %>% dplyr::select(-MEQ,MEQ)
  }
  names(data)[ncol(data)] <- 'y'
  for (i in (1:(dim(data)[2] -1))) {
    res = glm(y ~ .,data=subset(data, select=c(i, ncol(data))),family=binomial())
    a <- c( colnames(data)[i], exp(summary(res)$coef[2,][1]), exp(confint(res)[2,]), summary(res)$coef[2,][4])
    df[nrow(df) + 1,] = a
  }
  adjusted <- p.adjust(df[, c('P_Value')], method="BH")
  df$Adjusted_P = adjusted
  df <- df[order(df$P_Value),]
  
  df
}
```

##Multivariate
```{r, results='hide', include = FALSE}
multivariate_lr <- function(gender, disease) {

  options(scipen=999)
  df <- tibble(Variable_Name=character(), Odds_Ratio=numeric(), Conf_25=numeric(), Conf_75=numeric(), P_Value=numeric())
  
  f_path <- paste(disease,'/combined/',gender,'/combined.sav',sep='')
  data2 <- read.spss(file =f_path, to.data.frame = T)
  if (grepl('BDI',disease,fixed=TRUE)) {
    y <- as.formula(BDI ~ .)
  } else if (grepl('STAI',disease,fixed=TRUE)) {
    y <- as.formula(STAI ~ .)
  } else if (grepl('SPAQ',disease,fixed=TRUE)){
    y <- as.formula(SPAQ ~ .)
  } else if (grepl('SLEEP',disease,fixed=TRUE)){
    y <- as.formula(SLEEP ~ .)
  } else {
    y <- as.formula(MEQ ~ .)
  }
  
  multi = glm(y, data=data2, family=binomial())
  trm <- attr(multi$terms, "term.labels")

  
  for (i in seq(2, dim(data2)[2], by=1)){

    if (is.na(multi$coefficients[i])) {
      df <- df %>% add_row(Variable_Name = trm[i-1], 
                           Odds_Ratio = NA, 
                           Conf_25 = NA, 
                           Conf_75 = NA,
                           P_Value = NA)
    } else {
      a <- tryCatch(exp(confint(multi, trm[i-1])), error=function(e) c(NA,NA))
      b <- tryCatch(exp(summary(multi)$coef[i,][1]), error=function(e) c(NA))
      c <- tryCatch(summary(multi)$coef[i,][4], error=function(e) c(NA))
      df <- df %>% add_row(Variable_Name = trm[i-1], 
                           Odds_Ratio = b, 
                           Conf_25 = a[1],  
                           Conf_75 = a[2],
                           P_Value = c)
    }
  }

  adjusted2 <- p.adjust(as.numeric(unlist(df %>% dplyr::select(P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted2)
  
  df <- df %>% arrange(P_Value)
  df
}
```

```{r, results='hide', include = FALSE}

multivariate_lr_top <- function(disease) {
  f_path = paste(disease, ".csv", sep ="")
  data2 <- data.frame(read.csv(f_path))
  data2 = data2[,!names(data2) %in% c("X")]
  options(scipen=999)
  df <- tibble(Variable_Name=character(), Odds_Ratio=numeric(), Conf_25=numeric(), Conf_75=numeric(), P_Value=numeric())
  
  if (grepl('BDI',disease,fixed=TRUE)) {
    y <- as.formula(BDI ~ .)
  } else if (grepl('STAI',disease,fixed=TRUE)) {
    y <- as.formula(STAI ~ .)
  } else if (grepl('SPAQ',disease,fixed=TRUE)){
    y <- as.formula(SPAQ ~ .)
  } else if (grepl('SLEEP',disease,fixed=TRUE)){
    y <- as.formula(SLEEP ~ .)
  } else {
    y <- as.formula(MEQ ~ .)
  }
  
  multi = glm(y, data=data2, family=binomial())
  trm <- attr(multi$terms, "term.labels")

  
  for (i in seq(2, dim(data2)[2], by=1)){

    if (is.na(multi$coefficients[i])) {
      df <- df %>% add_row(Variable_Name = trm[i-1], 
                           Odds_Ratio = NA, 
                           Conf_25 = NA, 
                           Conf_75 = NA,
                           P_Value = NA)
    } else {
      a <- tryCatch(exp(confint(multi, trm[i-1])), error=function(e) c(NA,NA))
      b <- tryCatch(exp(summary(multi)$coef[i,][1]), error=function(e) c(NA))
      c <- tryCatch(summary(multi)$coef[i,][4], error=function(e) c(NA))
      df <- df %>% add_row(Variable_Name = trm[i-1], 
                           Odds_Ratio = b, 
                           Conf_25 = a[1],  
                           Conf_75 = a[2],
                           P_Value = c)
    }
  }

  adjusted2 <- p.adjust(as.numeric(unlist(df %>% dplyr::select(P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted2)
  
  df <- df %>% arrange(P_Value)
  df
}

```

#BIC and AIC for top features
``` {r, results='hide', include = FALSE}
library(RcmdrMisc)
library(rcompanion)
library(MASS)
bicaictop = function(data,disease, IC){
  data = data
  options(scipen=999)
  if (grepl('BDI',disease,fixed=TRUE)) {
    y <- as.formula(BDI ~ .)
  } else if (grepl('STAI',disease,fixed=TRUE)) {
    y <- as.formula(STAI ~ .)
  } else if (grepl('SPAQ',disease,fixed=TRUE)){
    y <- as.formula(SPAQ ~ .)
  } else if (grepl('SLEEP',disease,fixed=TRUE)){
    y <- as.formula(SLEEP ~ .)
  } else {
    y <- as.formula(MEQ ~ .)
  }
  
  first.glm = glm(y, data = data, family = 'binomial')

  res = stepwise(first.glm, trace = F, direction = "forward/backward", criterion = IC)
  Variable_Name = rownames(summary(res)$coefficients)
  df = data.frame(Variable_Name)
  df <- cbind(df, coef(res))
  df <- cbind(df, confint(res))
  df <- cbind(df, P_Value = summary(res)$coefficients[,4])
  r.squared = rep((nagelkerke(res)$Pseudo.R.squared.for.model.vs.null)[3], nrow(df))
  df = cbind(df, r.squared)
  adjusted2 <- p.adjust(as.numeric(unlist(dplyr::select(df, P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted2)
  
  df <- df %>% arrange(P_Value)
  df
  
}
```

#Runs Multivariate Regression for top features 
``` {r, results='hide',warning = F, include = FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("splits/Top")
disease_vec = c("BDI14","BDI20", "BDI28", "STAI38", "STAI45", "SLEEP25","SLEEP29")# 
for (disease in disease_vec){
  f_path = paste(disease, ".csv", sep ="")
  data <- read.csv(f_path)
  data = data[,!names(data) %in% c("X")]
  df = bicaictop(data, disease, "BIC")
  fname = paste(paste("bictop", disease, sep = "_"), ".csv", sep = "")
  write.csv(df, file = fname, row.names = FALSE)
}
```

# Fisher Exact Values

```{r, results='hide', include = FALSE}

fisherexact <- function(gender, disease) {
  f_path <- paste(disease,'/combined/',gender,'/combined.sav',sep='')
  data <- read.spss(file =f_path, to.data.frame = T)
  data$MEQ[data$MEQ <=59] = 0
  data$MEQ[data$MEQ>59] = 1
  data = data[,(which(colSums(data) != 0))]
  data = data[, (which(colSums(data) != nrow(data)))] 
  #glimpse(data)
  df = tibble(P_Value=numeric())
  #glimpse(df)
  for (i in 1:(dim(data)[2]-1)){
    c <- crosstab(data[,dim(data)[2]], data[,i], fisher=TRUE)
    df <- df %>% add_row(P_Value=c$fisher.ts$p.value)
  }

  adjusted <- p.adjust(as.numeric(unlist(dplyr::select(df,P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted)
  
  names = colnames(data)[1:(dim(data)[2]-1)]
  df <- df %>% add_column(Variable_Name = names)

  df <- df %>% arrange(P_Value)
  df
}
```

```{r, include = FALSE}
create_table <- function(dataframe, label) {
  kableExtra::kable_styling(knitr::kable(dataframe[dataframe$P_Value <= 0.05,], booktabs = TRUE, 
      caption = label), font_size = 10)
}
```

# Bayesian and Akaike Information Criterion Logistic Regressions
```{r, results='hide', include = FALSE}
bicaicglm <- function(data, gender, disease, IC) {
  data = data
  options(scipen=999)
  
  if (grepl('BDI',disease,fixed=TRUE)) {
    y <- as.formula(BDI ~ .)
  } else if (grepl('STAI',disease,fixed=TRUE)) {
    y <- as.formula(STAI ~ .)
  } else if (grepl('SPAQ',disease,fixed=TRUE)){
    y <- as.formula(SPAQ ~ .)
  } else if (grepl('SLEEP',disease,fixed=TRUE)){
    y <- as.formula(SLEEP ~ .)
  } else {
    y <- as.formula(MEQ ~ .)
  }
  
  first.glm = glm(y, family = binomial, data = data)
  res = stepwise(first.glm, trace = F, direction = "forward/backward", criterion = IC)
  r.squared = rep((nagelkerke(res)$Pseudo.R.squared.for.model.vs.null)[3], nrow(df))
  df = cbind(df, r.squared)
  Variable_Name = rownames(summary(res)$coefficients)
  df = data.frame(Variable_Name)
  df <- cbind(df, coef(res))
  df <- cbind(df, confint(res))
  df <- cbind(df, P_Value = summary(res)$coefficients[,4])
  
  
  adjusted2 <- p.adjust(as.numeric(unlist(dplyr::select(df, P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted2)
  
  df <- df %>% arrange(P_Value)
  df
  
}
```

#Mediation Analysis -- Linear Regression
```{r, results='hide', warning= F, include = FALSE}
library(mediation)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("splits/LinReg") 
p_cutoff = 0.1
for (outcome in c("BDI", "SLEEP", "STAI")){
  f_path = paste(outcome, ".sav", sep ="")
  data =read.spss(f_path, to.data.frame = T)
  data = data[,(which(colSums(data) != 0))]
  data = data[, (which(colSums(data) != nrow(data)))] 
  mediator = "MEQ"
  columns = colnames(data)
  columns = columns[!columns %in% c(outcome,mediator)]
  for (col in columns){
    f1 = as.formula(paste(mediator,"~",col, sep =""))
    model.M = lm(f1, data = data)
    if (summary(model.M)$coef[,4][2] <=p_cutoff){
      f2 = as.formula(paste(outcome, "~", paste(mediator, col,sep= "+"),sep =""))
      model.Y = lm(f2, data = data)
      results = mediate(model.M, model.Y, treat = col, mediator =mediator,boot =T,
                      sims = 500)
      if (results$d.avg.p <= p_cutoff){
        f3 = as.formula(paste(outcome, "~", col, sep = ""))
        model.YX = lm(f3, data = data)
        if (summary(model.YX)$coef[,4][2] <=p_cutoff){
          sink(paste(paste(col, mediator,outcome, sep ="_"), ".txt", sep =""))
          print(summary(results))
          print(summary(model.M))
          print(summary(model.Y))
          print(summary(model.YX))
          sink()
          }
        }
      }
  }
  # formula = as.formula(paste("BDI ~", paste(CF, collapse= "+"), sep = ""))
  # model_CF_BDI = lm(formula, data =data)
  # summary(model_CF_BDI)
  # model_CF_MEQ = lm(as.formula(paste("MEQ ~", paste(CF, collapse= "+"), sep = "")), data =data)
  # summary(model_CF_MEQ)
  # model_all = lm(BDI~., data = data)
  # summary(model_all)
}
```

#This code runs the AIC and BIC Linear Regressions

```{r, results='hide', include = FALSE}
library(RcmdrMisc)
bicaiclm <- function(data, outcome, IC) {
  data =data
  if (grepl('BDI',outcome,fixed=TRUE)) {
    y <- as.formula(BDI ~ .)
  } else if (grepl('STAI',outcome,fixed=TRUE)) {
    y <- as.formula(STAI ~ .)
  } else if (grepl('SPAQ',outcome,fixed=TRUE)){
    y <- as.formula(SPAQ ~ .)
  } else if (grepl('SLEEP',outcome,fixed=TRUE)){
    y <- as.formula(SLEEP ~ .)
  }
  model_a = lm(y, data= data) 
  res = stepwise(model_a, direction = "forward/backward", criterion = IC)
  Variable_Name = rownames(summary(res)$coefficients)
  df = data.frame(Variable_Name)
  df <- cbind(df, Coef = summary(res)$coefficients[,1])
  df <- cbind(df, confint(res))
  df <- cbind(df, P_Value = summary(res)$coefficients[,4])
  df$r_squared <- summary(res)$r.squared
  df$adj_r_squared <- summary(res)$adj.r.squared
  
  adjusted2 <- p.adjust(as.numeric(unlist(dplyr::select(df, P_Value))), method="BH")
  df <- df %>% add_column(Adjusted_P = adjusted2)
  
  df <- df %>% arrange(P_Value)
  df
  
}
```

```{r, results='hide', include = FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("splits/LinReg")
for (outcome in c("BDI", "SLEEP", "STAI")){
  f_path = paste(outcome, ".sav", sep ="")
  data =read.spss(f_path, to.data.frame = T)
  df = bicaiclm(data, outcome, "BIC")
  fname = paste(paste("linreg_BIC", outcome, sep = "_"),".csv", sep ="")
  write.csv(df, file = fname, row.names = FALSE)
  
  df = bicaiclm(data, outcome, "AIC")
  fname = paste(paste("linreg_AIC", outcome, sep = "_"),".csv", sep ="")
  write.csv(df, file = fname, row.names = FALSE)
}
```



#Code that runs all statistical tests except Linear Regressions and ANOVAs
Change diseases depending on your cutoffs/outcomes. Change methods based on which
tests you want to use
```{r, results='hide', warning= FALSE, include = FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

disease_vec = c("BDI14","BDI20","BDI28", "STAI38", "STAI45", "SLEEP25", "SLEEP29") #"BDI", "STAI", "SLEEP", "SPAQ8", "SPAQ11"
clinicial_vec = c("cli") #"nocli"
ways_vec = c("2")
genders_vec = c("gendered")# "male", "female")#,
methods = c("univariate", "multivariate","fisher","bicglm", "aicglm") # 
# "univariate_top_10", "univariate_top_20","multivariate_top_10","multivariate_top_20" in case you want
#to perform these on top features.

setwd("splits")

for (method in methods){
  for (clinical in clinicial_vec){
    for (disease in disease_vec){
      for (ways in ways_vec){
        for (gender in genders_vec){
          outcome = paste(disease,"_", ways,clinical, sep = "")
          if (method == "univariate"){
            df = univariate_lr(gender, outcome)
          }
          else if (method == "univariate_top_10"){
            df = univariate_lr_top(outcome, 10)
          }
          else if (method == "univariate_top_20"){
            df = univariate_lr_top(outcome, 20)
          }
          else if (method == "multivariate"){
            df = multivariate_lr(gender, outcome)
          }
          else if (method == "multivariate_top"){
            df = multivariate_lr_top(disease)
          }
          else if (method == "multivariate_top_20"){
            df = univariate_lr_top(outcome, 20)
          }
          else if (method == "fisher"){
            df = fisherexact(gender, outcome)
          }
          else if(method == "bicglm"){
            f_path <- paste(outcome,'/combined/',gender,'/combined.sav',sep='')
            data <- read.spss(file =f_path, to.data.frame = T)
            
            df = bicaicglm(data,gender, outcome, 'BIC')
          }
          else {
            f_path <- paste(outcome,'/combined/',gender,'/combined.sav',sep='')
            data <- read.spss(file =f_path, to.data.frame = T)
            
            df = bicaicglm(data, gender, outcome, 'AIC')
          }
          fname = paste(paste(method, gender, outcome, sep = "_"), ".csv", sep = "")
          write.csv(df, file = fname, row.names = FALSE)
        }
      }
    }
  }
}
```


## ANOVA/Kruskal-Wallis/Shreirer-Ray-Hare
```{r, results='hide', warning= FALSE, include = FALSE}
library(car)
library(rcompanion)
library(FSA)
library(foreign)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
setwd("splits/anova")
aov_data = read.spss("anova.sav", to.data.frame = T)
aov_data$MEQ[aov_data$MEQ >59] = "MT"
aov_data$MEQ[aov_data$MEQ<41] = "ET"
aov_data$MEQ[aov_data$MEQ>=41 & aov_data$MEQ <=59] = "IT"

aov_data$AGE[aov_data$AGE<=22] = 1
aov_data$AGE[aov_data$AGE>22] = 0
BDI = aov_data$BDI
SLEEP = aov_data$SLEEP
STAI = aov_data$STAI

outcomes = c("BDI", "SLEEP", "STAI")
aov_data = subset(aov_data, select = -c(MSF, SPAQ, BDI, SLEEP, STAI))
aov_data = data.frame(lapply(aov_data, as.factor))
aov_data = apply(aov_data, 2, function(x) gsub("^\\s*$", NA, x))
aov_data = data.frame(aov_data)
columns = colnames(aov_data)
num_cols = ncol(aov_data) 
options(contrasts = c(unordered="contr.sum", ordered="contr.poly"))
for (outcome in outcomes) {
  for (i in 1:(num_cols-1)){
    for (j in (i+1):num_cols){
      small_df = aov_data[, c(i,j)]
      small_df = data.frame(lapply(small_df, as.factor))
      if (outcome == "BDI") {
        small_df = cbind(small_df, y = BDI)
      }
      else if (outcome == "SLEEP") {
        small_df = cbind(small_df, y = SLEEP)
      } else {
        small_df = cbind(small_df, y = STAI)
      }
      
      small_df = na.omit(small_df)
      small_df$y = rank(small_df$y)
      formula = as.formula(paste("y~", paste(colnames(small_df)[1:2], collapse = "*"), sep = ""))
      anova_a = lm(formula, data = small_df)
      anova_2 = Anova(anova_a, type = 3, singular.ok = T)
      df = data.frame(Names = rownames(anova_2)[-1])
      df = cbind(df, Sum_Sq = anova_2$`Sum Sq`[-1])
      df = cbind(df, degf = anova_2$Df[-1])
      MSTOT = sum(anova_2$`Sum Sq`)/sum(anova_2$Df)
      df = cbind(df, H = (anova_2$`Sum Sq`[-1])/MSTOT)
      chi_vals = 1 - pchisq(df$H,df$degf)
      df = cbind(df, P_Value = chi_vals)
      if (sum(chi_vals < 0.1)>0 & nrow(df) ==4){
        fname = paste(paste(paste(colnames(small_df)[1:2],collapse = "_"), outcome, sep ="_"),
                      ".csv", sep ="")
        write.csv(df, file = fname, row.names = F)
        if (nlevels(small_df[,1])>2) {
          write.csv(dunnTest(small_df[,'y'], small_df[,1], method = 'bh')$res,
                      file = paste("Dunn1_", fname,sep = ""),
                      row.names = F)
        }
        if (nlevels(small_df[,2])>2){
          write.csv(dunnTest(small_df[,'y'], small_df[,2], method = 'bh')$res, 
                      file = paste("Dunn2_", fname,sep = ""),
                      row.names = F)
        }
      }
    }
  }
    
  
}
```


## Association Rule Learning!

``` {r, warning = F, include = F}
library(arules)
library(wordcloud)
library(arulesViz)
library(jpeg)
library(tiff)
library(igraph)
library(dplyr)
library(Rgraphviz)
library(tidygraph)
library(ggraph)
library(stringr)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
outcomes = c('STAI')

for (outcome in outcomes){
  fname = paste(outcome, "_SMOTE_data", ".csv", sep ="")
  data = read.csv(fname)
  data = data[,!names(data) %in% c("X")]# "AGE", "GENDER", "SOCIOSTATUS"
  data <- transform(data, MEQ = ifelse(MEQ <= 41, "ET", ifelse(MEQ >= 59,"MT","IT")))
  data = data %>% mutate(MEQ_ET = ifelse(MEQ == "ET",1,0), MEQ_MT = ifelse(
    MEQ == "MT", 1,0)) %>% dplyr::select(-MEQ)
  #data = transform(data, MEQ = ifelse(MEQ <=41, 1, 0))
  data <- transform(data, AGE = ifelse(AGE <=22, 1, 0))
  #data <- transform(data, GENDER = ifelse(GENDER ==1, "Female", "Male"))
  data_F = subset(data, GENDER ==1)
  data_M = subset(data, GENDER == 0)
  data_F = dplyr::select(data_F, -GENDER)
  data_M = dplyr::select(data_M, -GENDER)
  for (sex in c("Female", "Male")){
    if (sex == "Female"){data_set = data_F}
    else {data_set = data_M}
    data_set <- data.frame(lapply(data_set,as.logical))
    rhs_rule = paste(outcome, sep="")
    rules <- apriori(data_set, parameter = list(supp=0.04, conf=0.5,maxlen=10,target ="rules"),appearance = list(default="lhs",rhs=rhs_rule), control=list(verbose = FALSE, load = T, memopt = F))
    filtered_rules = subset(rules, subset = confidence >0.7)
    filtered_rules = arules::sort(filtered_rules, decreasing = TRUE, na.last = NA, 
      by = c("confidence", "lift", "support"), order = FALSE)
    f_out =paste("arules_", outcome, "_", sex, ".csv", sep ="")
    arules::write(filtered_rules, file = f_out, sep = ",")
    
    lhs = labels(lhs(filtered_rules))
    rules.df = DATAFRAME(filtered_rules)
    lift.vals = rules.df$lift
    lhs = gsub('[{}]', '', lhs)
    list.of.rules = strsplit(lhs, ",")
    lhs = paste(lhs, collapse = ",")
    lhs = unlist(strsplit(lhs, ","))
    LHS = data.frame(lhs = lhs)
    frequency_table = count(LHS, lhs)
    
    tiff(filename = paste("wordcloud_", sex, ".tiff", sep =""), width =8, height =8,
         units = "in", res =300)
    wordcloud(words = frequency_table$lhs, freq = frequency_table$n, min.freq = 1,
              colors=brewer.pal(8, "Dark2"),random.order=FALSE, rot.per=0.4,
              scale = c(1, 3), random.color = F, main = sex)
    dev.off()
    # top_12 = head(filtered_rules, n = 12, by ="lift")
    # jpeg(width = 1100, height= 600, res =100,
    #      filename = "association_graph.jpeg")
    # plot(top_12, method = "graph", control =list(type = "items"))
    # dev.off()
    
    feature_names = frequency_table$lhs
    count_matrix = matrix(data = 0, nrow = length(feature_names),
                          ncol = length(feature_names))
    
    colnames(count_matrix) = feature_names
    rownames(count_matrix) = feature_names
    
    lift_total = data.frame(Feature = feature_names,
                            Total_lift = rep(0, length(feature_names)),
                            Count = frequency_table$n)
    for (i in 1:length(list.of.rules)){
      temp.vec = list.of.rules[[i]]
      #print(temp.vec)
      if (length(temp.vec) != 1){
        for (feature in temp.vec){
          lift_total$Total_lift[which(lift_total$Feature ==feature)] = lift_total$Total_lift[which(lift_total$Feature ==feature)] + lift.vals[i]
        }
        for (j in 1:(length(temp.vec)-1)){
          for (k in (j+1):length(temp.vec)){
            count_matrix[temp.vec[j], temp.vec[k]] = 
              count_matrix[temp.vec[j], temp.vec[k]] + 1
          }
        }
      }
    }
    lift_total = lift_total %>% mutate(Average_Lift = Total_lift/Count)
    count_matrix[count_matrix <10] = 0
    to.remove = (colSums(count_matrix) ==0 &
                                    rowSums(count_matrix) ==0)
    lift_total = lift_total %>%
   filter(!row_number() %in% which(to.remove))
    count_matrix = count_matrix[!(colSums(count_matrix) ==0 &
                                    rowSums(count_matrix) ==0),
                                !(colSums(count_matrix) ==0 &
                                    rowSums(count_matrix) ==0)]
    matrix_names = colnames(count_matrix)
    matrix_names = str_replace(matrix_names, "SOCIOSTATUS_3.0", "Upper Middle Class")
    if ("SOCIOSTATUS_4.0" %in% matrix_names){
      matrix_names = str_replace(matrix_names, "SOCIOSTATUS_4.0", "Upper Class")}
    matrix_names = str_replace(matrix_names, "_", "-")
    colnames(count_matrix) = matrix_names
    rownames(count_matrix) = matrix_names
    graph = igraph::graph_from_adjacency_matrix(count_matrix, mode = "undirected",
                                        diag = F, weighted = T)
    
    
    
    
    sz = lift_total$Average_Lift
    width = round(proportions(E(graph)$weight)*100,3)
    # new_width = case_when(width > quantile(width, 0.75) ~ , sz < quantile(sz, 0.25) ~ "red",
    #                TRUE ~ "yellow"))
    
    if(sex == "Female"){new_sz = case_when(sz > quantile(sz, 0.75) ~ "green", sz < quantile(sz, 0.25) ~ "red",
                   TRUE ~ "yellow")}
    else {new_sz = case_when(sz > quantile(sz, 0.75) ~ "black",
                             sz < quantile(sz, 0.25) ~ "purple",
                   TRUE ~ "blue")}
    if (sex == "Female"){colors_vec = c("green", "yellow", "red")}
    else {colors_vec = c("black", "purple", "blue")}
    
    if (sex == "Female"){
      F_graph = graph
      F_graph_names = V(graph)$name
      }
    else {
      M_graph = graph
      M_graph_names = V(graph)$name
      }
    
    
    
    tiff(filename = paste("graph_", sex, ".tiff", sep =""), width =9, height =8,
         units = "in", res =300)
    coords = layout_(F_graph, as_star())
    if (sex == "Male"){coords = coords[match(M_graph_names, F_graph_names),]}
    plot(graph, layout = coords,
         edge.width = exp(proportions(E(graph)$weight)*30), 
         normalize = F,
         edge.arrow.mode = 0, vertex.color = new_sz, vertex.size =proportions(lift_total$Count)*50,
         vertex.label.cex = 1.2, vertex.label.font = 2, vertex.label.color= "black",
         vertex.label.dist = case_when(
           coords[,1] <0 ~-1.9,
           coords[,2] == 0 & coords[,1]>0 ~ 2.85,
           coords[,1] >0 ~ 1.7,
           TRUE ~ 1.9),
         vertex.label.degree = case_when(
           coords[,2]<0 & coords[,1]<0  ~ -pi/4,
           coords[,2] > 0 & coords[,1] > 0 ~ -pi/4,
           coords[,2] == 0 & coords[,1] > 0 ~ -pi/4 +0.2,
           #coords[,2] < 0 & coords[,1] < 0 ~ pi/4,
           TRUE ~ pi/4 + 0.1)
         , main = sex)
    dummy_leg = legend("topleft", legend = c(paste(">",round(quantile(sz, 0.75),2),sep =""),
                            paste(round(quantile(sz, 0.25),2), " - ",
                                  round(quantile(sz, 0.75),2),
                                  sep =""),
                            paste("<",round(quantile(sz, 0.25),2),sep ="")),
           col = colors_vec, pch = c(19,19,19),cex =0.8,
           title = "Average Lift", plot = F)
    legend("topleft", legend = c(paste(">",round(quantile(sz, 0.75),2),sep =""),
                            paste(round(quantile(sz, 0.25),2), " - ",
                                  round(quantile(sz, 0.75),2),
                                  sep =""),
                            paste("<",round(quantile(sz, 0.25),2),sep ="")),
           col = colors_vec, pch = c(19,19,19),cex =1.1,
           title = "Average Lift")
    # legend("bottomright",
    #        legend = paste0(c(round(quantile(width, 0.75)),
    #                         round(quantile(width, 0.50)),
    #                         round(quantile(width, 0.25))), "%"),
    #        col = "grey",
    #        lwd = c(quantile(exp(proportions(E(graph)$weight)*30),
    #                         c(0.75,0.5,0.25))),
    #        cex = 1.1,
    #        title = "Co-occurence")    
    dev.off()
    
  }

  
  
  # g1 = plot(as(count_matrix, "graphNEL"))
  # 
  # tidy_graph = as_tbl_graph(count_matrix)
  # g2= ggraph(tidy_graph, layout = "tree") + 
  # geom_node_point(aes(color = weight), alpha =) +
  # geom_edge_link(aes(width = weight), alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 2)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "Frequency") +
  # theme_graph()
  # g2
}



```

#Code for bar graph figures
``` {r, warning = F, results = 'hide', include = F, message = F}
library(foreign)
library(ggplot2)
library(BSDA)
library(tidyverse)
library(gridExtra)
library(rcompanion)
library(car)
library(emmeans)


create_bargraph <- function(gene1, gene2, SNP1,SNP2, Mood) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
  setwd("splits/anova")
  data = read.spss("anova.sav", to.data.frame = T)
  data_subset = data[,names(data) %in% c(gene1, gene2, "GENDER", Mood)]
  data_subset = apply(data_subset, 2, function(x) gsub("^\\s*$", NA, x))
  data_subset = data.frame(data_subset)
  data_subset = na.omit(data_subset)
  data_subset[,Mood] = as.numeric(data_subset[, Mood])
  data_subset$GENDER = as.numeric(data_subset$GENDER)
  data_subset$GENDER = ifelse(data_subset$GENDER ==2, "Female", "Male")
  colnames(data_subset)[which(names(data_subset) == "GENDER")] <- "Gender"
  gene_combination =paste(gene1, "-", gene2, sep = "")
  genotype_name = paste(SNP1, "-", SNP2, sep ="")
  Genotype = ifelse(data_subset[,gene1] == SNP1 & data_subset[,gene2] == SNP2, genotype_name, "Others")
  data_subset = cbind(data_subset, Genotype)
  data_subset$Genotype = factor(Genotype, levels = c(genotype_name, "Others"))
  
  options(contrasts = c(unordered="contr.sum", ordered="contr.poly"))
  
  sink(file =  paste(gene1, gene2, SNP1,SNP2, Mood, ".txt", sep =""))
  subset_anova = aov(STAI ~ Gender*Genotype, data = data_subset)
  subset_anova_3 = Anova(subset_anova, type = 3)
  print(subset_anova_3, quote = T)
  print(shapiro.test(subset_anova$residuals))
  print(hist(subset_anova$residuals))
  print(leveneTest(STAI ~ Gender*Genotype, data = data_subset))
  print(paste("n =", nrow(data_subset)))
  
  TukeyGender = emmeans(subset_anova,~Gender)
  print(pairs(TukeyGender, adjust="tukey"))
  TukeyGenotype = emmeans(subset_anova,~Genotype)
  print(pairs(TukeyGenotype, adjust="tukey"))
  TukeyInt = emmeans(subset_anova,~Gender*Genotype)
  print(pairs(TukeyInt, adjust="tukey"))
    
  sink()
  
  summary = summarize(group_by(data_subset, Genotype), Average = mean(!!sym(Mood)), SE = (sd(!!sym(Mood))/sqrt(length(!!sym(Mood)))))
  print(summary)
  plot_1 = ggplot(summary, aes(Genotype, Average, fill = Genotype)) +
  geom_col() + geom_bar(stat = "identity", color = 'black')+
  geom_errorbar(aes(ymin = Average - SE, ymax = Average +SE), width=0.2) +
    labs(y=paste("Average ",Mood," Score",sep=""), x = gene1) + theme_classic() + 
    scale_fill_manual(values=c("#FFFFFF", "#808080"))+coord_cartesian(ylim= c(20,50))
  
  
  summary_gendered = summarize(group_by(data_subset, Genotype, Gender),
                               Average = mean(!!sym(Mood)) -20, SE = (sd(!!sym(Mood))/sqrt(length(!!sym(Mood)))))
  plot_2 = ggplot(summary_gendered, aes(Gender, Average, fill = Genotype)) +
    geom_bar(stat = "identity", color = 'black',
                                          width=0.85, position=position_dodge(0.95))+
  geom_errorbar(aes(ymin = Average - SE, ymax = Average +SE), width=0.2, position = position_dodge(0.95)) +labs(y=paste("Average ",Mood," Score",sep=""), x = "Gender", fill = gene_combination) + theme_classic() + scale_y_continuous(expand = c(0,0), limits = c(0, 40.5),breaks=seq(0,40,10), labels=seq(20,60,10))+
    scale_fill_manual(values=c("#FFFFFF", "#808080"))+
    theme(
      axis.text=element_text(size=17), axis.title = element_text(size = 18),
      legend.title = element_text(size= 18), legend.text = element_text(size =17))
  
  
  
  
  summary = summarize(group_by(data_subset, !!sym(gene1),!!sym(gene2)),
                      Average = mean(!!sym(Mood)), SE = (sd(!!sym(Mood))/sqrt(length(!!sym(Mood)))))
  plot_3 = ggplot(summary, aes(!!sym(gene1), Average, fill = !!sym(gene2))) +
  geom_col(position = position_dodge(0.95)) + geom_bar(stat = "identity", color = 'black',
                                          width=0.9, position=position_dodge(0.95))+
  geom_errorbar(aes(ymin = Average - SE, ymax = Average +SE), width=0.2, position = position_dodge(0.95)) +labs(y=paste("Average ",Mood," Score",sep=""), x = gene1) + theme_classic() + 
    scale_fill_manual(values=c("#FFFFFF", "#808080", '#D3D3D3'))+coord_cartesian(ylim= c(20,60))
  
  summary_gendered_gene = summarize(group_by(data_subset, !!sym(gene1),!!sym(gene2), Gender), Average = mean(!!sym(Mood)) - 20, SE = (sd(!!sym(Mood))/sqrt(length(!!sym(Mood)))))
  plot_4 = ggplot(summary_gendered_gene, aes(!!sym(gene1), Average, fill = !!sym(gene2)))+geom_bar(stat = "identity", color = 'black',
                                          width=0.7, position=position_dodge(width = 0.9, preserve = "single"))+
  geom_errorbar(aes(ymin = Average - SE, ymax = Average +SE), width=0.2, position = position_dodge(width = 0.9, preserve = 'single')) +labs(y=paste("Average ",Mood," Score",sep=""), x = gene1) + theme_classic() + 
    scale_y_continuous(expand = c(0,0), limits = c(0, 41),breaks=seq(0,40,10), labels=seq(20,60,10))+
    scale_fill_manual(values=c("#FFFFFF", '#D3D3D3',"#808080"))+ facet_wrap(~ Gender)+theme(
      axis.text=element_text(size=17), axis.title = element_text(size = 18),
      legend.title = element_text(size= 18), legend.text = element_text(size =17),
      strip.text.x = element_text(size = 18))+
    geom_hline(yintercept = 0)
  
  tiff(filename = paste(gene1, gene2, SNP1,SNP2, Mood, "1",".tiff", sep =""),
       width =8, height =8,
         units = "in", res =300)  
  print(plot_2)
  dev.off()
  tiff(filename = paste(gene1, gene2, SNP1,SNP2, Mood, "2",".tiff", sep =""),
       width =8, height =8,
         units = "in", res =300)  
  print(plot_4)
  dev.off()
  pdf(file = paste(gene1, gene2, SNP1,SNP2, Mood, ".pdf", sep =""), onefile = T)
  print(summary_gendered)
  print(summary_gendered_gene)
  print(plot_1)
  print(plot_2)
  print(plot_3)
  print(plot_4)
  dev.off()
    
  plot_4
  
}

```

#Code that uses the above function to make bar graphs
``` {r, warning = F, results = 'hide', include = F}
#change based on the genotypes provided by Dr. Ingram

create_bargraph("CLOCK3111", "CRY2", "TC", "AG", "STAI")
create_bargraph("CLOCK3111", "CRY2", "TC", "AG", "SLEEP")
create_bargraph("PER3B", "CRY2", "AG", "AA", "STAI")
create_bargraph("CLOCK3111", "CRY1", "TT", "CC", "SLEEP")
create_bargraph("PER2", "VNTR", "AA", "BB", "BDI")
create_bargraph("PER3B", "CRY1", "AG", "CC", "SLEEP")
create_bargraph("PER3B", "CRY1", "AG", "CG", "BDI")
create_bargraph("PER3B", "CRY1", "AG", "CG", "BDI")
create_bargraph("PER3B", "CRY1", "AG", "CG", "STAI")
create_bargraph("PER3B", "CRY1", "AG", "CG", "SLEEP")
create_bargraph("CRY1", "VNTR", "CC", "AA", "STAI")
create_bargraph("PER2", "PER3B", "AA", "AG", "SLEEP")
create_bargraph("PER3A", "CRY1", "GG", "GG", "BDI")
```